The Role of Context in Spatial Region Identification

### 2013-09-13 code/scenarios/cdsr/trunk/papers/ACS2013/main.tex

\author{Kate Lockwood}{klockwood@csumb.edu}
\address{ITCD Department, California State University - Monterey Bay}
\author{John D. Kelleher}{john.d.kelleher@dit.ie}
\address{Applied Intelligence Research Centre, Dublin Institute of Technology }
\author{Matthew Klenk}{klenk@parc.com}
\address{Palo Alto Research Center, Palo Alto CA }
\author{Nick Hawes}{n.a.hawes@cs.bham.ac.uk}
\address{Intelligent Robotics Lab, University of Birmingham, UK }
\vskip 0.2in

\begin{abstract}
Representing and reasoning about spatial terms is an essential ability for cognitive systems interacting with humans in a shared environment.  Consider an agent asked to remain ``in the safe zone'' during a military engagement.  The size, shape, and location of this area depend on the capabilities of the agent, other agents in the scene, and the constraints imposed by the environment.  We call these types of regions \textit{context-dependent spatial regions} (CDSRs) due to the context-dependent nature of their extent.  We claim that understanding these regions requires integrating semantic and geometric knowledge about a dynamic environment as well as understanding how task-specific information influences spatial interpretation.  We report the results of a human study conducted to explore how people interpret three types of CDSRs and to support the knowledge integration claim.  We discss the implications for these results on cognitive systems development.
\end{abstract}

\section{Introduction} 
Natural collaboration between humans and cognitive systems requires both participants to communicate using a shared understanding of spatial language.  Cognitive systems can already understand spatial terms defined in terms of coordinates on a map~\cite{Zender2008a}, or that can be specified using known landmarks~\cite{brenneretal07ijcai}.  However, many spatial terms are defined not only by their geometry, but also their context and their relevance to current goals.  That is, to understand this language requires considering the other objects and agents in the environment, as well as their configuration and functional roles.  We call regions that can only be understood in this way \textit{context-dependent spatial regions} (CDSRs).

We claim that (1) people can effectively communicate using these regions, and (2) the size, location and shape of these regions change with the context.  We argue that to account for this behavior, a cognitive systems solution must respond to changing context, apply learned regions to new contexts, and reuse learned knowledge in multiple tasks.  To explore these claims, we present the results of a human study, in which, people marked regions and rated points representing CDSRs in three different environments in 13 different contexts.  In discussing the results, we consider a number of approaches from current cognitive systems and robotics.  Each approach has a number of shortcomings that must be addressed not only to improve our understanding of human spatial reasoning, but also to enable robust intelligent assistants in human environments.

@@

Consider the following regions (shown in Figure \ref{fig:examples}): the front of a classroom, safety in a military engagement, and a geographic bottleneck.  To determine these regions, one must identify the physical boundaries (e.g., walls, bodies of water), the objects within the environment, their orientation and their functional use (e.g., desks oriented toward a whiteboard), and other agents' intentions and capabilities (e.g., enemy tanks can move and have a range of attack).%  Furthermore, these regions exist at different levels of spatial resolution.  Consider a neighborhood in a city, whose boundaries expand and contract over time with changes to the surrounding residents and businesses.

@@

Understanding CDSRs is crucial to joint intelligent action with humans; among other reasons they are frequently tieds to goals and commands: ``go to the front of the classroom'', ``defend the bottleneck'', and ``get to safety''. In the rest of the paper we present a study that examines the contextual dependence of these regions.  Following this we discuss the implications of these regions for the development of cognitive systems.

@@

Both of these research questions address the fundamental claim that CDSRs are contextually defined. Question 1 addresses this claim by examining: (a) whether or not a CDSR is always applicable in a given enviroment (e.g. does a classroom always have a front?), and (b) the variation in CDSR placement and size within different contexts. Research question 2 addresses the contextual nature of CDSRs in terms of whether or not the variation in the location and extent of a CDSR systematically follow the variations  in the given context.

@@

We developed two experimental tasks that each address both research questions:
\begin{itemize}
	\item{Task 1:} subjects were presented with a visual representation of a spatial scene plus a linguistic description containing a CDSR and were tasked with drawing a polygon on the visual stimuli that defined the CDSR (or stated there was no such region).
	\item{Task 2:} subjects were presented with a visual representation of a spatial scene with a marked location and an accompanying spatial description and were asked to rate how well the location fit with their understanding of the CDSR on a five point likert scale.
\end{itemize}

@@

The classroom stimuli consisted of pairs of floor plans and photos of classrooms with different chair configurations. For example, Figure \ref{fig:classroom-a} is a floor plan of a classroom where the chairs have been arranged in a circle and Figure \ref{fig:classroom-b} is the accompanying photo of this classroom. We developed four different classroom stimuli each with a different layout of desks: a circular layout, a U shaped layout (Figure \ref{fig:classroom-c} and Figure \ref{fig:classroom-d}), a typical row-based layout with the first row of chairs close to the expected front of the room (Figure \ref{fig:classroom-e}), and the same arrangement with the front row of chairs further away from the expected front of the room (Figure \ref{fig:classroom-f})\footnote{In the interest of saving space we do not show all the classroom photo stimuli}. To control for left-right bias effects, two versions of the U shaped layout, close rows layout and far rows layout were created by horizontally flipping the images. Consequently, there were seven classroom stimuli (circle, U $\times$ 2, close $\times$ 2, far $\times$ 2).  The two geographical map stimuli, Figures \ref{fig:bottleneck-a} and \ref{fig:bottleneck-b}, differed in the extent of the water.  Finally, the four military engagement stimuli, Figures \ref{fig:tanks-a}, \ref{fig:tanks-b}, \ref{fig:turrets-a}, \ref{fig:tanks-b} varied the functionality of the objects in the scene (tanks that could move versus fixed turret emplacements), the location of the objects in the scenes (were the tanks close to the bottom of the image or near the top), and the number and layout of the objects (the number and layout of the turrets were varied).

@@

\subsection{Task 1 - Design}
In task one, the subject was asked to draw a polygon that defined the extent of the CDSR given or to select a no region option.  The drawing was done on the diagram in all stimuli with the classroom photographs provided for reference in those situations.  Subjects drew polygons by creating a sequence of vertices with lines being automatically generated between them as the subject drew.  There was no upper bound placed on the number of line segments drawn but subjects were required to provide at least three line segments or to choose no region before continuing.

@@

We use \textbf{Cohen's kappa coefficient}, $\kappa$, to measure the agreement between two regions as follows.  We generated a set of 47,000 test points that were uniformly distributed across each stimuli and used each subject's polygon to classify the test points as being inside or outside the polygon. After classifying the test points for each subject we then computed an average pairwise kappa score for each subject with all the other subjects on a per stimuli basis. In this calculation if either of the subjects in a pair entered a no region response for a particular stimuli a $\kappa$ value of 0 was used.  $\kappa$ ranges from 0 indicating no agreement beyond chance to 1 indicating complete agreement.  Typically, scores above 0.4 are considered fair agreement and 0.75 considered excellent.

@@

%However, this calculation was done across all the stimuli and as we have already mentioned some of the stimuli were deliberately designed to be non-prototypical, in particular we expected the classroom stimuli where the chairs were arranged in a circle (Figure \ref{fig:classroom-a}) and in a U shaped pattern (Figure \ref{fig:classroom-c} and its horizontally flipped equivalent) to be relatively novel contexts for the subjects. Indeed, if we look at the histogram of the average pairwise $\kappa$ scores by stimuli, Figure \ref{fig:stim-kappa}, it is evident that there is a large variation, ranging from $0$ to $0.8$, in $\kappa$ scores for each stimuli.

\begin{table}[t]
\parbox{.50\linewidth}{
\centering 
\begin{tabular}{l c c}
% Stim ID & Average Pairwise $\kappa$ & No Region Responses\\
 & Average & No \\
Stim ID & Pairwise $\kappa$ & Reg.\\
\hline
\hline
(h) & 0.5034 & 0\\ % 1_2
(g) & 0.4585 & 1\\ % 1_1
(k) & 0.4548 & 0\\ % 3_3
(f) & 0.3930 & 0\\ % 2_4
(i) & 0.3906 & 0\\ % 3_1
(e-flipped) & 0.3813 & 2\\ % 2_3
(j) & 0.3787 & 1\\ % 3_2
(f-flipped) & 0.3720 & 2\\ % 2_5
(e) & 0.3629 & 1\\ %2_2
(l) & 0.3164& 0\\ % 3_4
(c) & 0.2438 & 2\\ % 2_6
(c-flipped) & 0.2414 & 4\\ %2_7
(a) & 0.0996 & 10 \\ % 2_1
\hline
\end{tabular}
\caption{The average pairwise $\kappa$ scores and the number of no region response by visual stimuli.}
\label{tab:stimuli-kappa-scores}
}
\hfill
\parbox{.50\linewidth}{
\centering
\begin{tabular}{c c}
Expected Relationship & Human Results  \\
\hline
\hline
area(g) > area(h) & 291.2 > 276.2* \\
maxX(f) > maxX(e) & 176.1 > 167.1 \\
% minimumX(e-flipped) > minimumX(f-flipped) & 336.1 > 321.3 (p<.12) \\
minX(e-flipped)  &  \\
 > minX(f-flipped) & 336.1 > 321.3* \\
maxY(i) > maxY(j) & 304.64 > 265.04* \\
maxY(j) > maxY(k) & 265.04 > 251.71 \\
maxY(k) = maxY(l) & 265.04 != 290.03 \\
\hline
\end{tabular}
\caption{Comparison of the expected relationships to the mean human responses.}
\label{tab:region-location-tests}
}
\end{table}

@@

Table \ref{tab:stimuli-kappa-scores} lists the average pairwise $\kappa$ score for each of the visual stimuli used in the experiment and the number of no region responses for each stimuli.  It is clear from this table that the less common stimuli have the lowest $\kappa$ scores.  Each of the non-typical regions, the turret +1 arrangement (l), the u-shaped layouts (c and c-flpped), and the circle layout have kappa scores of less than 0.35. Looking at the classroom results as a whole, these results support the hypothesis that the definition of CDSRs in a given context is effected by the location and orientation of objects in the environment.  This dependence is observed by the fact that just moving the desks into a circle configuration the CDSR for the front of the room no longer exists.

@@

Table \ref{tab:region-location-tests} compares the expected relationships between the geometric extent of a CDSR in different contexts.  In five of the six comparisons, the expected relationships were found, supporting our claim that changes of context systematically change the perceived extent of a CDSR.  The comparison of the safety area with tanks (j) versus turrets (k) is interesting as it provides some evidence that people are simulating actions of other agents in the scene.  Because the tanks can move and turrets cannot, the safe region extends higher on screen in the tanks context.  The case that did not demonstrate the expected the relationship was the safety region between the row of turrets (k) and turrets +1(l).   A possible explanation for this is that there is a domain expertise component to understanding contexts.

@@

\subsection{Task 2 - Design} 
In task two, a location was marked on the stimuli (using a red dot) and the subject was asked to rank on a five point likert scale how well the location matched the given CDSR, or select not applicable.  The location of these points was informed by the results of a pilot study, such that one of the domain's points, called the \textbf{sweetspot}, was located near the expected prototypical center of the CDSR. In the classroom domain, the three points were located along the horizontal line that bisected the stimuli with P1 positioned on the left of the stimuli, P2 in the center and P3 on the right of the stimuli. This provides comparable data points between the trials where the classroom stimuli were horizontally flipped.  In the geography situations, P1 was positioned in the top left corner of the stimuli, P2 was the expected sweetspot and was located within the expected bottleneck region for both stimuli and P3 was located in the bottom right corner of the stimuli (see Figure \ref{fig:bottleneck-points}).  In the military engagement domain the three test points were located on the vertical line that bisected the stimuli: P1 was located near the enemy tanks, P3 was located with all friendly units between the it and the enemy tanks and P2 was located near the center of the image as follows. In stimuli (j,k) the point lay between all forces, in stimuli (i), it lay behind the friendly tanks, and in stimuli (l) , it behind the lone turret but in front of the row turrets.  P3 was the expected sweetspot in the military engagement domain but the rating of P2 should differ with context receiving the highest score in (i) and the lowest score in (k).

@@

Another way of testing the relationship between the variation in context and the variation in definition of the CDSRs is to look at the differences in the ranking of the P2 point in the military domain between the stimuli (i) and (j) and the stimuli (k) and (l). In stimuli (i) the friendly tanks are positioned between the P2 point and the enemy tanks whereas in stimuli (j) the P2 point is in no-mans land. Given this change in context one would expect a higher ranking for the P2 point in context (i) than in context (j). As the results in Table \ref{tab:point-likert-scores} show this is the case with this higher rating statiscially significant at a $95\%$ confidence level. Stimuli (k) and (l) present an analagous context with P2 being protected by an extra turret in stimuli (l) but in no-mans land in stimuli (k). Again, the average rating of P2 for stimuli (l) is higher than that of (k), although this difference is not statistically significant. Notice, that the rating of the sweetspot for the stimuli (i) and (l) remain quite high. This would indicate that the increase in the rating of the P2 points for these stimuli results from an increase in the area of the CDSR rather than a moving of the locus of the CDSR. Overall, the analysis of the ratings of the different sweetspots across the non-flipped and flipped stimuli and the variation in the ratings of the P2 points in the military domain stimuli point to the fact that both the location and the extent of a CDSR is contextually dependent.

@@

Taken together these results indicate that: (a) for all stimuli there was a location that was considered to be inside the region described by the linguistic term and that other locations were considered to be outside of this region; (b) within each domain the particular location that was considered to be the sweetspot was determined by the particular context depicted in the visual stimuli (i.e., changes in the position and orientation of objects with functional significance effected the location of the region); and (c) the impact of the changes in the depicted context was consistent across similar contexts, as shown by the fact that the sweetspot flipped from P1 to P3 when the context was flipped.

@@

Spatial reasoning has been an active topic in AI and robotics for decades, CDSRs present a new problem which requires solutions beyond the current state-of-the-art. Regions are central to QSR work~\cite{Cohn:2001} and also in the area of cognitive robitics (see \emph{inter alia.} ~\cite{brenneretal07ijcai,kelleher/costello:09}). A fundamental decision when creating a computational model of a CDSR is the method used to represent its boundary. Given their context-dependent nature, boundary definitions based purely on single geometric shapes or a fixed number of points, seem unlikely to model the regions with sufficient accuracy. However, arbitrary boundaries can be defined using a sequence of geometric primitives (usually straight lines). This is an approach we followed in previous work \cite{Hawes:2012}, where we defined CDSRs using \textit{anchor points} as the vertices of the region polygon thereby linking the conceptual regions to perceived objects.  While satisfying the requirements of a cognitive systems solution, applying this model to the results of our experiment illustrates two limitations.  First, Regions are not always physically bounded by objects.  In the classroom examples, people frequently did not extend the front of the room all way to the desks.  Second, people consider dynamics of a context, in the safety example, people did not indicate that it was safe behind the lone turret in front of the others. Finally, the connection between spatial regions and function has been explored in other robotics work, e.g. ~\cite{Karg:2012,Fasola:2013}, but these regions are based on fields anchored at single points in the environment or on a particular object. Therefore, they cannot account for the changes in context shown in this experiment which depend on multiple objects, agents, and simulation.  McLure and Forbus \cite{McLureForbus:2012} present a learning system that selects which encoding schemes work best for identifying geographic features (e.g., bottlenecks).  While this approach incorporates functional aspects of the situation, it is unclear how it would apply to the other situations in our experiment.

@@

\section{Discussion}
Cognitive systems do not face problems in isolation.  Context, including its goals, objects in the environment, and other agents, impact how information is processed and how actions are undertaken.  In this paper, we presented a human study to develop requirements for reasoning about context dependent spatial regions for cognitive systems.  We found that (1) people can communicate with CDSRs, although their ability to do so decreases when situations become ambiguous, and (2) varying the location and types of the objects and agents in an environment produced changes in the location, size, and shape of the CDSR.  We outline requirements for a cognitive systems solution and discuss limitations in current approaches from cognitive systems and robotics.  Solutions to this problem will significantly improve the interaction between people and intelligent systems in human environments with applications to domestic robots, security, and activity recognition.

>>>>

\author{Kate Lockwood}{klockwood@csumb.edu}
\address{ITCD Department, California State University - Monterey Bay}
\author{John D. Kelleher}{john.d.kelleher@dit.ie}
\address{Applied Intelligence Research Centre, Dublin Institute of Technology }
\author{Matthew Klenk}{klenk@parc.com}
\address{Palo Alto Research Center, Palo Alto CA }
\author{Nick Hawes}{n.a.hawes@cs.bham.ac.uk}
\address{School of Computer Science, University of Birmingham, UK }
\vskip 0.2in

\begin{abstract}
Representing and reasoning about spatial terms is an essential ability for cognitive systems interacting with humans in a shared environment.  Consider an agent asked to remain ``in the safe zone'' during a military engagement.  The size, shape, and location of this area depend on the capabilities of the agent, other agents in the scene, and the constraints imposed by the environment.  We call these types of regions \textit{context-dependent spatial regions} (CDSRs) due to the context-dependent nature of their extent.  We claim that understanding these regions requires integrating semantic and geometric knowledge about a dynamic environment as well as understanding how task-specific information influences spatial interpretation.  We report the results of a human study conducted to explore how people interpret three types of CDSRs and discuss the implications for developing a cognitive systems with this ability.
\end{abstract}

\section{Introduction} 
Natural collaboration between humans and cognitive systems requires both participants to communicate using a shared understanding of spatial language.  Cognitive systems can already understand spatial terms defined in terms of coordinates on a map~\cite{Zender2008a}, or that can be specified using known landmarks~\cite{brenneretal07ijcai}.  However, many spatial terms are defined not only by their geometry, but also their context.  That is, to understand this language requires considering the other objects and agents in the environment, as well as their configuration, functional roles, and goals.  We call regions that can only be understood in this way \textit{context-dependent spatial regions} (CDSRs).

We claim that (1) people can effectively communicate using these regions, and (2) the size, location and shape of these regions change with the context.  To explore these claims, we present the results of a human study, in which, people marked regions and rated points representing CDSRs in three different environments in 13 different contexts.  In discussing the results, we consider a number of approaches from current cognitive systems and robotics.  We posit that to account for this ability, a cognitive system must respond to changing context, apply learned regions to new contexts, and reuse learned knowledge in multiple tasks.  Each approach has a number of shortcomings that must be addressed not only to improve our understanding of human spatial reasoning, but also to enable robust intelligent assistants in human environments.

@@

Consider the following regions (shown in Figure \ref{fig:examples}): the front of a classroom, safety in a military engagement, and a geographic bottleneck.  To determine the extent of these regions, one must identify the physical boundaries (e.g., walls, bodies of water), the objects within the environment, their orientation and their functional use (e.g., desks oriented toward a whiteboard), and other agents' intentions and capabilities (e.g., enemy tanks can move and have a range of attack).%  Furthermore, these regions exist at different levels of spatial resolution.  Consider a neighborhood in a city, whose boundaries expand and contract over time with changes to the surrounding residents and businesses.

@@

Understanding CDSRs is crucial to joint intelligent action with humans; among other reasons they are frequently tieds to goals and commands: ``go to the front of the classroom'', ``defend the bottleneck'', and ``get to safety''. In the rest of the paper we present a study that examines the contextual dependence of these regions.  Following this we discuss how the results align with previous research and requirements for a cognitive systems model of this ability.

@@

Both of these research questions address the fundamental claim that CDSRs are contextually defined. Question 1 addresses this claim by examining: (a) whether or not a CDSR is always applicable in a given enviroment (e.g. does a classroom always have a front?), and (b) the variation in CDSR placement and size within different contexts. Question 2 addresses the contextual nature of CDSRs in terms of whether or not the variation in the location and extent of a CDSR systematically follow the variations  in the given context.

@@

We developed two experimental tasks that each address both research questions:
\begin{itemize}
	\item{Task 1:} Subjects were presented with a visual representation of a spatial scene plus a linguistic description containing a CDSR and were tasked with drawing a polygon on the visual stimuli that defined the CDSR (or stating there was no such region).
	\item{Task 2:} Subjects were presented with a visual representation of a spatial scene with a marked location and an accompanying spatial description and were asked to rate how well the location fit with their understanding of the CDSR on a five point likert scale.
\end{itemize}

@@

The classroom stimuli consisted of pairs of floor plans and photos of classrooms with different chair configurations. For example, Figure \ref{fig:classroom-a} is a floor plan of a classroom where the chairs have been arranged in a circle and Figure \ref{fig:classroom-b} is the accompanying photo of this classroom. We developed four different classroom stimuli each with a different layout of desks: a circular layout, a U shaped layout (Figure \ref{fig:classroom-c} and Figure \ref{fig:classroom-d}), a typical row-based layout with the first row of chairs close to the expected front of the room (Figure \ref{fig:classroom-e}), and the same arrangement with the front row of chairs further away from the expected front of the room (Figure \ref{fig:classroom-f})\footnote{In the interest of saving space we do not show all the classroom photo stimuli}. To control for left-right bias effects, two versions of the U shaped layout, close rows layout and far rows layout were created by horizontally flipping the images. Consequently, there were seven classroom stimuli (circle, U $\times$ 2, close $\times$ 2, far $\times$ 2).  The two geographical map stimuli, Figures \ref{fig:bottleneck-a} and \ref{fig:bottleneck-b}, differed in the extent of the water.  Finally, the four military engagement stimuli, Figures \ref{fig:tanks-a}, \ref{fig:tanks-b}, \ref{fig:turrets-a}, \ref{fig:tanks-b} varied the functionality of the objects in the scene (tanks that could move versus fixed turret emplacements), the location of the objects in the scenes (tanks close to the bottom of the image versus near the top), and the layout of the objects (turrets in a row versus turrets in a row with a lone turret in front).

@@

\subsection{Task 1 - Design}
In task one, the subject was asked to draw a polygon that defined the extent of the CDSR given or to select a "no region" option.  The drawing was done on the diagram in all stimuli with the classroom photographs provided for reference in those situations.  Subjects drew polygons by creating a sequence of vertices with lines being automatically generated between them as the subject drew.  There was no upper bound placed on the number of line segments drawn but subjects were required to provide at least three line segments or to choose no region before continuing.

@@

We use \textbf{Cohen's kappa coefficient}, $\kappa$, to measure the agreement between two regions as follows.  We generated a set of 47,000 test points that were uniformly distributed across each stimuli and used each subject's polygon to classify the test points as being inside or outside the polygon. After classifying the test points for each subject, we then computed an average pairwise kappa score for each subject with all the other subjects on a per stimuli basis. In this calculation, if either of the subjects in a pair entered a no region response for a particular stimuli a $\kappa$ value of 0 was used.  $\kappa$ ranges from 0 indicating no agreement to 1 indicating complete agreement.  Typically, scores above 0.4 are considered fair agreement and 0.75 considered excellent.

@@

%However, this calculation was done across all the stimuli and as we have already mentioned some of the stimuli were deliberately designed to be non-prototypical, in particular we expected the classroom stimuli where the chairs were arranged in a circle (Figure \ref{fig:classroom-a}) and in a U shaped pattern (Figure \ref{fig:classroom-c} and its horizontally flipped equivalent) to be relatively novel contexts for the subjects. Indeed, if we look at the histogram of the average pairwise $\kappa$ scores by stimuli, Figure \ref{fig:stim-kappa}, it is evident that there is a large variation, ranging from $0$ to $0.8$, in $\kappa$ scores for each stimuli. 
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
%\newcommand\footnoteref[1]{\protected@xdef\@thefnmark{\ref{#1}}\@footnotemark}

\begin{table}[t]
\parbox{.50\linewidth}{
\centering 
\begin{tabular}{l c c}
% Stim ID & Average Pairwise $\kappa$ & No Region Responses\\
 & Average & No \\
Stim ID & Pairwise $\kappa$ & Reg.\\
\hline
\hline
(h) & 0.5034 & 0\\ % 1_2
(g) & 0.4585 & 1\\ % 1_1
(k) & 0.4548 & 0\\ % 3_3
(f) & 0.3930 & 0\\ % 2_4
(i) & 0.3906 & 0\\ % 3_1
(e-flipped) & 0.3813 & 2\\ % 2_3
(j) & 0.3787 & 1\\ % 3_2
(f-flipped) & 0.3720 & 2\\ % 2_5
(e) & 0.3629 & 1\\ %2_2
(l) & 0.3164& 0\\ % 3_4
(c) & 0.2438 & 2\\ % 2_6
(c-flipped) & 0.2414 & 4\\ %2_7
(a) & 0.0996 & 10 \\ % 2_1
\hline
\end{tabular}
\caption{The average pairwise $\kappa$ scores and the number of no region response by visual stimuli.}
\label{tab:stimuli-kappa-scores}
}
\hfill
\parbox{.50\linewidth}{
\centering
\begin{tabular}{c c}
Expected Relationship & Human Results  \\
\hline
\hline
area(g) > area(h) & 291.2 > 276.2\footnotemark[2] \\
maxX(f) > maxX(e) & 176.1 > 167.1 \\
% minimumX(e-flipped) > minimumX(f-flipped) & 336.1 > 321.3 (p<.12) \\
minX(e-flipped)  &  \\
 > minX(f-flipped) & 336.1 > 321.3\footnotemark[2] \\
maxY(i) > maxY(j) & 304.64 > 265.04\footnotemark[2] \\
maxY(j) > maxY(k) & 265.04 > 251.71 \\
maxY(k) = maxY(l) & 265.04 != 290.03 \\
\hline
\end{tabular}
\caption{Comparison of the expected relationships to the mean human responses.}
\label{tab:region-location-tests}
}
\end{table}

\footnotetext[2]{Significance value from T-Test (p<0.1)}

@@

Table \ref{tab:stimuli-kappa-scores} lists the average pairwise $\kappa$ score for each of the visual stimuli used in the experiment and the number of no region responses for each stimuli.  It is clear from this table that the less common stimuli have the lowest $\kappa$ scores.  Each of the non-typical regions, the turret +1 arrangement (l), the u-shaped layouts (c and c-flpped), and the circle layout have kappa scores of less than 0.35. Looking at the classroom results as a whole supports the hypothesis that the definition of CDSRs in a given context is effected by the location and orientation of objects in the environment.  This dependence is observed by the fact that just moving the desks into a circle configuration the shared understanding of the front of the room no longer exists.

@@

Table \ref{tab:region-location-tests} compares the expected relationships between the geometric extent of a CDSR in different contexts.  In five of the six comparisons, the expected relationships were found, supporting our claim that changes of context systematically change the perceived extent of a CDSR.  The comparison of the safety area with tanks (j) versus turrets (k) is interesting as it provides some evidence that people are simulating actions of other agents in the scene.  Because the tanks can move and turrets cannot, the safe region extends higher on screen in the tanks context.  The case that did not demonstrate the expected the relationship was the safety region between the row of turrets (k) and turrets +1(l).  That is, a single turret does not provide safety from three tanks.  A possible explanation for this is that there is a domain expertise component to understanding contexts.

@@

\subsection{Task 2 - Design} 
In task two, a location was marked on the stimuli (using a red dot) and the subject was asked to rank on a five point likert scale how well the location matched the given CDSR, or select not applicable.  The location of these points was informed by the results of a pilot study, such that one of the domain's points, called the \textbf{sweetspot}, was located near the expected prototypical center of the CDSR. In the classroom domain, the three points were located along the horizontal line that bisected the stimuli with P1 positioned on the left of the stimuli, P2 in the center and P3 on the right of the stimuli. This provides comparable data points between the trials where the classroom stimuli were horizontally flipped.  In the geography situations, P1 was positioned in the top left corner of the stimuli, P2 was the expected sweetspot and was located within the expected bottleneck region for both stimuli and P3 was located in the bottom right corner of the stimuli.  In the military engagement domain the three test points were located on the vertical line that bisected the stimuli: P1 was located near the enemy tanks, P3 was located with all friendly units between the it and the enemy tanks and P2 was located near the center of the image as follows. In stimuli (j,k) the point lay between all forces, in stimuli (i), it lay behind the friendly tanks, and in stimuli (l) , it lay behind the lone turret but in front of the row turrets.  P3 was the expected sweetspot in the military engagement domain but the rating of P2 should differ with (i) receiving the highest score and (k) the lowest score.

@@

Another way of testing the relationship between the variation in context and the variation in definition of the CDSRs is to look at the differences in the ranking of the P2 point in the military domain between the stimuli (i) and (j) and the stimuli (k) and (l). In stimuli (i) the friendly tanks are positioned between the P2 point and the enemy tanks whereas in stimuli (j) the P2 point is in no-mans land. Given this change in context one would expect a higher ranking for the P2 point in context (i) than in context (j). This is confirmed by the results in Table \ref{tab:point-likert-scores} and the higher rating is statistically significant $(p<.05)$. Stimuli (k) and (l) present an analagous context with P2 being protected by an extra turret in stimuli (l) but in no-mans land in stimuli (k). Again, the average rating of P2 for stimuli (l) is higher than that of (k), although this difference is not statistically significant. Thus, it is unclear the extent to which particiapnts were simulating the outcome of the battle. Notice, that the rating of the sweetspot for the stimuli (i) and (l) remain quite high. This would indicate that the increase in the rating of the P2 points for these stimuli results from an increase in the area of the CDSR rather than a moving of the locus of the CDSR. Overall, the analysis of the ratings of the different sweetspots across the non-flipped and flipped stimuli and the variation in the ratings of the P2 points in the military domain stimuli point to the fact that both the location and the extent of a CDSR is contextually dependent.

@@

Taken together these results indicate that: (1) for all stimuli there was a location that was considered to be inside the region described by the linguistic term and that other locations were considered to be outside of this region; (2) within each domain the particular location that was considered to be the sweetspot was determined by the particular context depicted in the visual stimuli (i.e., changes in the position and orientation of objects with functional significance effected the location of the region); and (3) the impact of the changes in the depicted context was consistent across similar contexts, as shown by the fact that the sweetspot flipped from P1 to P3 when the context was flipped.

@@

Cognitive systems seek to reproduce the full range of human behavior.  In addition to reproducing these results, a cognitive systems solution must learn about new regions and apply this knowledge in multiple tasks.  While spatial reasoning is an established area in AI and robotics, CDSRs present a new problem which requires solutions beyond the current state-of-the-art. Regions are central to QSR work~\cite{Cohn:2001} and also in the area of cognitive robitics (see \emph{inter alia.} ~\cite{brenneretal07ijcai,kelleher/costello:09}). A fundamental decision when creating a computational model of a CDSR is the representation of its boundary. Given their context-dependent nature, boundary definitions based purely on single geometric shapes or a fixed number of points, seem unlikely to model the regions with sufficient accuracy. However, arbitrary boundaries can be defined using a sequence of geometric primitives (usually straight lines). This is an approach we followed in previous work \cite{Hawes:2012}, where we defined CDSRs using \textit{anchor points} as the vertices of the region polygon thereby linking the conceptual regions to perceived objects. One limitation of the anchor point approach is that regions are not always physically bounded by objects.  In the classroom examples, people frequently did not extend the front of the room all way to the desks. The connection between spatial regions and function has been explored in other robotics work, e.g. ~\cite{Karg:2012,Fasola:2013}, but these regions are based on fields anchored at single points in the environment or on a particular object. Therefore, they cannot account for the changes in context shown in this experiment which depend on multiple objects, agents, and simulation.  McLure and Forbus \cite{McLureForbus:2012} present a learning system that selects which encoding schemes work best for identifying geographic features (e.g., bottlenecks).  While this approach incorporates functional aspects of the geography, it is unclear how it would apply to the other situations in our experiment.

@@

\section{Discussion}
Cognitive systems do not face problems in isolation.  Context, including its goals, objects in the environment, and other agents, impact how information is processed and how actions are undertaken.  In this paper, we presented a human study to develop requirements for reasoning about context dependent spatial regions for cognitive systems.  We found that (1) people can communicate with CDSRs, although their ability to do so decreases when situations become ambiguous, and (2) varying the location and types of the objects and agents in an environment produced changes in the location, size, and shape of the CDSR.  We posit that a cognitive system should be able to learn new regions, account for changing contexts, and apply this knoweldge in multiple reasoning tasks. Solutions to this problem will significantly improve the interaction between people and intelligent systems in human environments with applications to domestic robots, security, and activity recognition.
